A Horrocks-type theorem for even orthogonal $K_2$

### 2020-04-04 dl-paper.tex
Many minor fixes

\begin{df} \label{defP0}
Let $M$ be an ideal of a local ring $A$. Denote by $\widetilde{P}_{\alpha, M}(0)$ the subgroup of $\overline{\St}(\Phi, A[X], M[X])$ generated by the following 5 families of elements parameterized by $f \in M[X]$, $\xi \in A[X]$:
\begin{enumerate} 
 \item $z_{\beta}(Xf, \xi),\ \beta \in \Phi \text{ such that }\alpha + \beta \in \Phi\text{ or } \alpha - \beta \in \Phi,\ f \in M[X],\ \xi \in A[X];$
 \item $z_{\beta}(f, X\xi),\ \alpha - \beta \in \Phi;$
 \item $z_{\beta}(f, \xi),\ \beta \perp \alpha;$
 \item $x_{-\alpha}(X^2f)$;
 \item $x_{\alpha}(f)$. \end{enumerate}

@@

\begin{lemma}\label{P0_normal} The subgroup $\widetilde{P}_{\alpha, M}(0)$ is normal in $\widetilde{P}_{\alpha, M}(*)$. Moreover, there is a short exact sequence of groups, which is split by the map $m \mapsto x_{-\alpha}(mX)$ (we denote by $(M, +)$ the additive group of the ideal $M$):
\[\xymatrix{1 \ar[r] & \widetilde{P}_{\alpha,M}(0) \ar[r] & \widetilde{P}_{\alpha,M}(*) \ar@{->>}[r]^{p_\alpha} & (M, +) \ar@<5pt>@{-->}[l] \ar[r] & 1}.\] \end{lemma}
\begin{proof} We need to verify that the conjugate by $x_{-\alpha}(mX)$ to every generator $g$ of $\widetilde{P}_\alpha(0)$ listed in~Definition~\ref{defP0} belongs to $\widetilde{P}_\alpha(0)$.
The assertion is obvious for the generators of type 3 and 4.

@@

\begin{rem}\label{rem:c-DS} By~\eqref{Cdef} one has $c_\beta(f, X\xi) = x_{\beta} (f) \cdot z_{\beta}(-f, -X\xi)$. This implies that for $\beta \in Z_+(\alpha)$ (resp. $\beta \in Z_0(\alpha)$) the element $c_\beta(f, X\xi)$ is a product of two generators of type 2 (resp. 3).
Thus, the elements $c_{\beta}(f, X\xi)$ lie in $\widetilde{P}_{\alpha, M}(0)$ for all $\beta \in Z(\alpha)$.

It is easy to check that $\widetilde{P}_{\alpha, M}(0)$ contains the image of $Z_\alpha(A, M)$ under the natural embedding of $\overline{\St}(\Phi, A, M) \hookrightarrow \overline{\St}(\Phi, A[X], M[X])$.
In particular, if $M$ is the maximal ideal of $A$, the subgroup $\widetilde{P}_{\alpha, M}(0)$ contains relative Dennis--Stein and Steinberg symbols. It is also easy to see that the image of $\widetilde{P}_{\alpha, M}(0)$ under the homomorphism $ev_{X=0}^*$ coincides with $Z_\alpha(A, M)$. \end{rem}

The following lemma shows that $\widetilde{P}_{\alpha, M}(*)$ is sufficiently large.
\begin{lemma} \label{Pstar-large} Suppose that $(A, M)$ is a local pair. 
Then the subgroup $\widetilde{P}_{\alpha, M}(*)$ contains the subgroup $K(A[X], M[X]) \leq \overline{\St}(\Phi, A[X], M[X])$ defined in the beginning of~\cref{sec:computationOfK}. \end{lemma}
\begin{proof} Clearly, $\widetilde{P}_{\alpha, M}(*)$ contains the elements $x_\beta(Xf)$ for all $\beta \in \Phi$ and $z_\beta(Xf, \xi)$ for $\beta \in \Phi \setminus \{\pm \alpha\}$, $f \in M[X]$, $\xi\in A[X]$, therefore by~\cref{thm:Stepanov} $\widetilde{P}_{\alpha, M}(*)$ contains all of $\overline{\St}(\Phi, A[X], XM[X])$. The required assertion now follows from~\cref{Kgen-strong} and~\cref{rem:c-DS}. \end{proof}

\begin{rem} \label{Pstar-char} 
By the above lemma and~\cref{rem:c-DS} the subgroup $\widetilde{P}_{\alpha, M}(*)$ coincides with the preimage of the subgroup $Z_\alpha(A, M) \leq \overline{\St}(\Phi, A, M)$ under the homomorphism
 \[ev_{X=0}^*\colon \overline{\St}(\Phi, A[X], M[X]) \to \overline{\St}(\Phi, A, M).\] \end{rem}

@@

\begin{lemma} \label{P0-conj-invariant} Suppose $(A, M)$ is a local pair. Then for any $\beta \in Z(\alpha)$ and $b \in A$ the subgroups $\widetilde{P}_{\alpha, M}(0)$ and $\widetilde{P}_{\alpha, M}(*)$ are stable under conjugation by $x_\beta(b)$. \end{lemma}
\begin{proof}
Notice that both $Z_\alpha(A, M)$ and $K(A[X], M[X])$ are stable under the specified conjugation, which implies the assertion for $\widetilde{P}_{\alpha, M}(*)$. 
To obtain the assertion for $P_{\alpha, M}(0)$ consider the following commutative diagram. \[ \xymatrix{ \widetilde{P}_{\alpha, M}(0) \ar@{^{(}->}[r] \ar@{^{(}->}[d] & \widetilde{P}_{\alpha, M}(*) \ar@{->>}[r]^{p_{\alpha, M}} \ar@{^{(}->}[d] & (M, +) \ar@{^{(}->}[d] \\ \widetilde{P}_{\alpha, A}(0) \ar@{^{(}->}[r] & \widetilde{P}_{\alpha, A}(*) \ar@{->>}[r]^{p_{\alpha, A}} & (A, +)} \]
Notice that $x_\beta(b) \in \widetilde{P}_{\alpha, A}(0)$ therefore for $g \in \widetilde{P}_{\alpha, M}(0)$ one has \[p_{\alpha, M}(x_\beta(b) \cdot g \cdot x_\beta(-b)) = p_{\alpha, A}(x_\beta(b) \cdot g \cdot x_\beta(-b)) = p_{\alpha, A}(g) = 0,\] which implies the assertion.\end{proof}

For the rest of this section $M$ denotes the maximal ideal of $A$.
\begin{df}\label{S-def}
Define the subgroup $P_\alpha(0)$ (resp. $P_{\alpha}(*)$) of $\overline{\St}(\Phi, R, M[X, X^{-1}])$
 as the image of the subgroup $\widetilde{P}_{\alpha, M}(0)$ (resp. $\widetilde{P}_{\alpha, M}(*)$) 
 under the obvious homomorphism  $j_+ \colon \overline{\St}(\Phi, A[X]) \to \overline{\St}(\Phi, R)$.

@@

\begin{corollary}\label{SR:additivity} For $g \in P_\alpha(m)$ one has $S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am) \in P(\alpha, \tfrac{m}{1 + am})$, moreover, \[ S_\alpha(a+b, g) = S_\alpha\left(b,\ S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am)\right)\cdot h_\alpha(1+am). \]\end{corollary} \begin{proof}
Fix an element $g \in P_\alpha(m)$ and write it $g = g_0 \cdot x_{-\alpha}(mX)$ for some $g_0 \in P_\alpha(0)$.
Direct computation shows that
\begin{align*}
 S_\alpha(a, g) \cdot h_{\alpha}^{-1}(1+am) = S_\alpha(a, g_0) \cdot S_\alpha(a, x_{-\alpha}(mX)) \cdot h_{\alpha}^{-1}(1+am)&\text{ by~\cref{lem:Smult}} \\\
 = S_\alpha(a, g_0) \cdot x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot\langle a, m \rangle &\text{ by~\cref{lem:Tulenbaev-formula}.}
\end{align*}
By \cref{P0_conj} and \cref{rem:c-DS} the subgroup $P_\alpha(0)$ contains $S_\alpha(a, g_0)$ and the symbol $\langle a, m \rangle$. Thus, the expression in the right-hand side of the above formula lies in $P_\alpha\left(\tfrac{m}{1+am}\right)$.

@@

\begin{externaltheorem} \label{thm:P1glueing}
 Let $\Phi$ be a root system of type $\rA_{\geq 4}, \rD_{\geq 5}$ or $\rE_{6,7,8}$.
 Suppose that the map $j_R \colon \St(\Phi, B) \to \St(\Phi, A[X, X\inv])$ is injective.  
 Then the map $j_-$ is injective and the following commutative square is pullback
 \begin{equation} \nonumber
  \xymatrix{ \St(\Phi, A) \ar[r] \ar[d] & \St(\Phi, A[X]) \ar[d]^{j_+} \\ \St(\Phi, A[X\inv]) \ar[r]^{j_-} & \St(\Phi, A[X, X\inv]).}
 \end{equation}
\end{externaltheorem}
\begin{proof} Since $j_- = j_R j_B^-$, where $j_B^-$ is as in~\cref{prop43}, it suffices to show that $j_B^-$ is injective.
 Set $v_0 = [1, 1, 1] \in \overline{V}_T$ and suppose $g \in \Ker(j_B^-)$.
 By~\cref{prop43} holds $v_0 = j_B^-(g) \cdot v_0 = [1, g, 1]$ therefore, by the definition of $\overline{V}_T$, $g = i_-(\gamma)$ for some $\gamma \in G_M^0$ such that $j_+i_+(\gamma) = 1$.
 Since $j_+i_+$ is injective, we conclude that $g = 1$.

>>>>

\begin{df} \label{defP0}
Let $M$ be an ideal of a local ring $A$. Denote by $\widetilde{P}_{\alpha, M}(0)$ the subgroup of $\overline{\St}(\Phi, A[X], M[X])$ generated by the following 5 families of elements parameterized by $f \in M[X]$, $\xi \in A[X]$:
\begin{enumerate} 
 \item $z_{\beta}(Xf, \xi),\ \beta \in \Phi \text{ such that }\alpha + \beta \in \Phi\text{ or } \alpha - \beta \in \Phi;$
 \item $z_{\beta}(f, X\xi),\ \alpha - \beta \in \Phi;$
 \item $z_{\beta}(f, \xi),\ \beta \perp \alpha;$
 \item $x_{-\alpha}(X^2f)$;
 \item $x_{\alpha}(f)$. \end{enumerate}

@@

\begin{lemma}\label{P0_normal} The subgroup $\widetilde{P}_{\alpha, M}(0)$ is normal in $\widetilde{P}_{\alpha, M}(*)$. Moreover, there is a short exact sequence of groups, which is split by the map $m \mapsto x_{-\alpha}(mX)$ (we denote by $(M, +)$ the additive group of the ideal $M$):
\[\xymatrix{1 \ar[r] & \widetilde{P}_{\alpha,M}(0) \ar[r] & \widetilde{P}_{\alpha,M}(*) \ar@{->>}[r]^{p_\alpha} & (M, +) \ar@<5pt>@{-->}[l] \ar[r] & 1}.\] \end{lemma}
\begin{proof} We need to verify that the conjugate by $x_{-\alpha}(mX)$ to every generator $g$ of $\widetilde{P}_\alpha(0)$ listed in~\cref{defP0} belongs to $\widetilde{P}_\alpha(0)$.
The assertion is obvious for the generators of type 3 and 4.

@@

\begin{rem}\label{rem:c} By~\eqref{Cdef} one has $c_\beta(f, X\xi) = x_{\beta} (f) \cdot z_{\beta}(-f, -X\xi)$. This implies that for $\beta \in Z_+(\alpha)$ (resp. $\beta \in Z_0(\alpha)$) the element $c_\beta(f, X\xi)$ is a product of two generators of type 2 (resp. 3).
Thus, the elements $c_{\beta}(f, X\xi)$ lie in $\widetilde{P}_{\alpha, M}(0)$ for all $\beta \in Z(\alpha)$. \end{rem}
\begin{rem}\label{rem:DS}
It is easy to check that $\widetilde{P}_{\alpha, M}(0)$ contains the image of $Z_\alpha(A, M)$ under the natural embedding of $\overline{\St}(\Phi, A, M) \hookrightarrow \overline{\St}(\Phi, A[X], M[X])$.
In particular, if $M$ is the maximal ideal of $A$, the subgroup $\widetilde{P}_{\alpha, M}(0)$ contains relative Dennis--Stein and Steinberg symbols. It is also easy to see that the image of $\widetilde{P}_{\alpha, M}(0)$ under the homomorphism $ev_{X=0}^*$ coincides with $Z_\alpha(A, M)$. \end{rem}

The following lemma shows that $\widetilde{P}_{\alpha, M}(*)$ is sufficiently large.
\begin{lemma} \label{Pstar-large} Suppose that $(A, M)$ is a local pair. 
Then the subgroup $\widetilde{P}_{\alpha, M}(*)$ contains the subgroup $K(A[X], M[X]) \leq \overline{\St}(\Phi, A[X], M[X])$ defined in the beginning of~\cref{sec:computationOfK}. \end{lemma}
\begin{proof} Clearly, $\widetilde{P}_{\alpha, M}(*)$ contains the elements $x_\beta(Xf)$ for all $\beta \in \Phi$ and $z_\beta(Xf, \xi)$ for $\beta \in \Phi \setminus \{\pm \alpha\}$, $f \in M[X]$, $\xi\in A[X]$, therefore by~\cref{thm:Stepanov} $\widetilde{P}_{\alpha, M}(*)$ contains all of $\overline{\St}(\Phi, A[X], XM[X])$. The required assertion now follows from~\cref{Kgen-strong} and~\cref{rem:c}. \end{proof}

\begin{rem} \label{Pstar-char} 
By the above lemma and~\cref{rem:DS} the subgroup $\widetilde{P}_{\alpha, M}(*)$ coincides with the preimage of the subgroup $Z_\alpha(A, M) \leq \overline{\St}(\Phi, A, M)$ under the homomorphism \[ev_{X=0}^*\colon \overline{\St}(\Phi, A[X], M[X]) \to \overline{\St}(\Phi, A, M).\] \end{rem}

@@

\begin{lemma} \label{P0-conj-invariant} Suppose $(A, M)$ is a local pair. Then for any $\beta \in Z(\alpha)$ and $b \in A$ the subgroups $\widetilde{P}_{\alpha, M}(0)$ and $\widetilde{P}_{\alpha, M}(*)$ are stable under conjugation by $x_\beta(b)$. \end{lemma}
\begin{proof}
Notice that both $Z_\alpha(A, M)$ and $K(A[X], M[X])$ are stable under the specified conjugation, which implies the assertion for $\widetilde{P}_{\alpha, M}(*)$. 
To obtain the assertion for $\widetilde{P}_{\alpha, M}(0)$ consider the following commutative diagram. \[ \xymatrix{ \widetilde{P}_{\alpha, M}(0) \ar@{^{(}->}[r] \ar@{^{(}->}[d] & \widetilde{P}_{\alpha, M}(*) \ar@{->>}[r]^{p_{\alpha, M}} \ar@{^{(}->}[d] & (M, +) \ar@{^{(}->}[d] \\ \widetilde{P}_{\alpha, A}(0) \ar@{^{(}->}[r] & \widetilde{P}_{\alpha, A}(*) \ar@{->>}[r]^{p_{\alpha, A}} & (A, +)} \]
Notice that $x_\beta(b) \in \widetilde{P}_{\alpha, A}(0)$, therefore for $g \in \widetilde{P}_{\alpha, M}(0)$ one has \[p_{\alpha, M}(x_\beta(b) \cdot g \cdot x_\beta(-b)) = p_{\alpha, A}(x_\beta(b) \cdot g \cdot x_\beta(-b)) = p_{\alpha, A}(g) = 0,\] which implies the assertion.\end{proof}

For the rest of this section $M$ denotes the maximal ideal of $A$.
\begin{df}\label{S-def}
Define the subgroup $P_\alpha(0)$ (resp. $P_{\alpha}(*)$) of $\overline{\St}(\Phi, R, M[X, X^{-1}])$
 as the image of the subgroup $\widetilde{P}_{\alpha, M}(0)$ (resp. $\widetilde{P}_{\alpha, M}(*)$) 
 under $j_+ \colon \St(\Phi, A[X]) \to \St(\Phi, R)$.

@@

\begin{corollary}\label{SR:additivity} For $g \in P_\alpha(m)$ one has $S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am) \in P(\alpha, \tfrac{m}{1 + am})$, moreover, \[ S_\alpha(a+b, g) = S_\alpha\left(b,\ S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am)\right)\cdot h_\alpha(1+am). \]\end{corollary} \begin{proof}
Fix an element $g \in P_\alpha(m)$ and write it $g = g_0 \cdot x_{-\alpha}(mX)$ for some $g_0 \in P_\alpha(0)$.
Direct computation shows that
\begin{align*}
 S_\alpha(a, g) \cdot h_{\alpha}^{-1}(1+am) = S_\alpha(a, g_0) \cdot S_\alpha(a, x_{-\alpha}(mX)) \cdot h_{\alpha}^{-1}(1+am)&\text{ by~\cref{lem:Smult}} \\\
 = S_\alpha(a, g_0) \cdot x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot\langle a, m \rangle &\text{ by~\cref{lem:Tulenbaev-formula}.}
\end{align*}
By \cref{P0_conj} and \cref{rem:DS} the subgroup $P_\alpha(0)$ contains $S_\alpha(a, g_0)$ and the symbol $\langle a, m \rangle$. Thus, the expression in the right-hand side of the above formula lies in $P_\alpha\left(\tfrac{m}{1+am}\right)$.

@@

\begin{externaltheorem} \label{thm:P1glueing}
 Let $\Phi$ be a root system of type $\rA_{\geq 4}, \rD_{\geq 5}$ or $\rE_{6,7,8}$.
 Suppose that the map $j_R \colon \St(\Phi, B) \to \St(\Phi, A[X, X\inv])$ is injective.  
 Then the map $j_-$ is injective and the following commutative square is pullback
 \begin{equation} \nonumber
  \xymatrix{ \St(\Phi, A) \ar[r] \ar[d] & \St(\Phi, A[X]) \ar[d]^{j_+} \\ \St(\Phi, A[X\inv]) \ar[r]^{j_-} & \St(\Phi, A[X, X\inv]).}
 \end{equation}
\end{externaltheorem}
\begin{proof} Since $j_- = j_R j_B^-$, where $j_B^-$ is as in~\cref{prop43}, it suffices to show that $j_B^-$ is injective.
 Set $v_0 = [1, 1, 1] \in \overline{V}_T$ and suppose $g \in \Ker(j_B^-)$.
 By~\cref{prop43} holds $v_0 = j_B^-(g) \cdot v_0 = [1, g, 1]$, therefore by the definition of $\overline{V}_T$, $g = i_-(\gamma)$ for some $\gamma \in G_M^0$ such that $j_+i_+(\gamma) = 1$.
 Since $j_+i_+$ is injective, we conclude that $g = 1$.
